---
title: "Business Intelligence Lab Submission Markdown"
author: "<Specify your name here>"
date: "<Specify the date when you submitted the lab>"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

+---------------------------------------------------+--------------------------------------------------------------------------------------------+
| **Student ID Numbers and Names of Group Members** | \<list one student name and ID per line; you should be between 2 and 5 members per group\> |
|                                                   |                                                                                            |
|                                                   | 1.122998-Starfield-Glenn Oloo                                                              |
|                                                   |                                                                                            |
|                                                   | 2.134782-Starfield-Yasmin Choma                                                            |
|                                                   |                                                                                            |
|                                                   | 3.134783-Starfield- Moses Mbugua                                                           |
+---------------------------------------------------+--------------------------------------------------------------------------------------------+
| **BBIT 4.2 Group**                                | \<either group A, B, or C\>                                                                |
+---------------------------------------------------+--------------------------------------------------------------------------------------------+
| **Course Code**                                   | BBT4206                                                                                    |
+---------------------------------------------------+--------------------------------------------------------------------------------------------+
| **Course Name**                                   | Business Intelligence II                                                                   |
+---------------------------------------------------+--------------------------------------------------------------------------------------------+
| **Program**                                       | Bachelor of Business Information Technology                                                |
+---------------------------------------------------+--------------------------------------------------------------------------------------------+
| **Semester Duration**                             | 21^st^ August 2023 to 28^th^ November 2023                                                 |
+---------------------------------------------------+--------------------------------------------------------------------------------------------+

# Setup Chunk

**Note:** the following KnitR options have been set as the global defaults:\
`knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE,                        collapse = FALSE, tidy = TRUE)`.

> output:\
> \
> github_document:\
> toc: yes\
> toc_depth: 4\
> fig_width: 6\
> fig_height: 4\
> df_print: default\
> \
> editor_options:\
> chunk_output_type: console

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE,
                      collapse = FALSE, tidy = TRUE)
```

# \<Provide an Appropriate Title Here\>

Describe the code chunk here:

```{r Your Second Code Chunk -Code used to load Student Perfomance dataset}

```

```{library(readxl)}
Student_Perfomance_Dataset <- readxl::read_excel("data/Student Perfomance Dataset.xlsx")

View(Student_Perfomance_Dataset)
```

```{r Your Second Code Chunk Used to check the number of Variables and Rows in the student Perfomance Dataset}
# Load the dataset from an Excel file
library(readxl)
Student_Perfomance_Dataset <- read_excel("data/Student Perfomance Dataset.xlsx")
```

Describe the next code chunk here:

```{r Your Third Code Chunk- Used to check the data types of the Variables which can either be Numeric or Character}
library(readr)
sapply(Student_Perfomance_Dataset, class)
```

## \<You Can Have a Sub-Title Here\>

```{r Your Fourth Code Chunk - used to see the frequency for the variables that are categorical}

Student_Perfomance_Dataset_freq <- Student_Perfomance_Dataset$mentor
cbind(frequency = table(Student_Perfomance_Dataset_freq),
      percentage = prop.table(table(Student_Perfomance_Dataset_freq)) * 100)

Student_Perfomance_Dataset_freq <- Student_Perfomance_Dataset$`Attendance Waiver Granted: 1 = Yes, 0 = No`
cbind(frequency = table(Student_Perfomance_Dataset_freq),
      percentage = prop.table(table(Student_Perfomance_Dataset_freq)) * 100)

```

```{r Your Third Code Chunk- Used to perform the mode, that is the variable that is repeated the most in Mentor "0" is the mode meaning most students do not have mentors}
library(readr)
Student_Perfomance_Dataset_mode <- names(table(Student_Perfomance_Dataset$mentor))[
  which(table(Student_Perfomance_Dataset$mentor) == max(table(Student_Perfomance_Dataset$mentor)))
]
print(Student_Perfomance_Dataset_mode)

Student_Perfomance_Dataset_mode <- names(table(Student_Perfomance_Dataset$`Attendance Waiver Granted: 1 = Yes, 0 = No`))[
  which(table(Student_Perfomance_Dataset$`Attendance Waiver Granted: 1 = Yes, 0 = No`) == max(table(Student_Perfomance_Dataset$`Attendance Waiver Granted: 1 = Yes, 0 = No`)))
]
print(Student_Perfomance_Dataset_mode)

```

```{r Your Third Code Chunk- Used to calculate the distribution of each variable that is Median, mean , 1st Quartile,3rd Quartile}
library(readr)
summary(Student_Perfomance_Dataset)
```

```{r Your Third Code Chunk- Used to calculate the the standard deviation for variables that data type are numeric}
library(readr)

numeric_columns <- Student_Perfomance_Dataset[, sapply(Student_Perfomance_Dataset, is.numeric)]
standard_deviations <- sapply(numeric_columns, sd)
print(standard_deviations)
```

```{r Your Third Code Chunk- Used to calculate the variance of each variable}
library(readr)
numeric_columns <- Student_Perfomance_Dataset[, sapply(Student_Perfomance_Dataset, is.numeric)]
variances <- sapply(numeric_columns, var)
print(variances)

```

```{r code chunk - Used to install kurosis packages and calculate Kurotis}
if (!is.element("e1071", installed.packages()[, 1])) {
  install.packages("e1071", dependencies = TRUE)
}
require("e1071")

sapply(Student_Perfomance_Dataset[, 26],  kurtosis, type = 2)
```

```{r code chunk - Used to calculate skweness}
sapply(Student_Perfomance_Dataset[, 26],  skewness, type = 2)
```

```{r code chunk - Used to calculate covarinace}
Student_Perfomance_Dataset_cov <- cov(Student_Perfomance_Dataset[, -27])
View(Student_Perfomance_Dataset_cov)
```

```{r code chunk - Used to calculate correlation}
Student_Perfomance_Dataset_cor <- cor(Student_Perfomance_Dataset [, c(1, 23, 26)])
View(Student_Perfomance_Dataset)
```

```{r code chunk - Used to calculate One-way ANOVA}
Student_Perfomance_Dataset_one_way_anova <- aov(gender ~ GRADE, data = Student_Perfomance_Dataset)
summary(Student_Perfomance_Dataset_one_way_anova)
```

```{r code chunk -  Used to display a layout of the histogram - step 17}

# Set the layout for the histograms
par(mfrow = c(1, ncol(Student_Perfomance_Dataset)))

# Loop through each numeric column and create a histogram
for (i in 1:ncol(Student_Perfomance_Dataset)) {
  if (is.numeric(Student_Perfomance_Dataset[, i])) {
    hist(Student_Perfomance_Dataset[, i], main = names(Student_Perfomance_Dataset)[i])
  }
}

```

```{r code chunk - Used to create Box and Whisker Plots - step 18}

# Create box and whisker plots for numeric attributes
par(mfrow = c(1, 3))
numeric_cols <- sapply(Student_Perfomance_Dataset, is.numeric)

for (i in 1:ncol(Student_Perfomance_Dataset)) {
  if (numeric_cols[i]) {
    boxplot(Student_Perfomance_Dataset[, i], main = names(Student_Perfomance_Dataset)[i])
  }
}
```

```{r code chunk - Used to create Bar Plots for Each Categorical Attribute - step 19}

# Load necessary libraries (if not already loaded)
if (!require(ggplot2)) {
  install.packages("ggplot2")
  library(ggplot2)
}

# Create a bar plot for "Absenteeism_percentage"
ggplot(Student_Perfomance_Dataset, aes(x = factor(round(`Absenteeism Percentage`, 2)))) +
  geom_bar() +
  labs(title = "Absenteeism Percentage", x = "Percentage", y = "Count") +
  theme_minimal()

# Create a bar plot for "Grade"
ggplot(Student_Perfomance_Dataset, aes(x = GRADE)) +
  geom_bar() +
  labs(title = "Grade Distribution", x = "GRADE", y = "Count") +
  theme_minimal()



```

```{r code chunk - Used to create a Missingness Map to Identify Missing Dat - step 20}


if (!is.element("Amelia", installed.packages()[, 1])) {
  install.packages("Amelia", dependencies = TRUE)
}
require("Amelia")

missmap(Student_Perfomance_Dataset, col = c("red", "grey"), legend = TRUE)
```

```{r code chunk - Used to create a Correlation Plot- step 21}

# Install and load the necessary packages
if (!is.element("corrplot", installed.packages()[, 1])) {
  install.packages("corrplot", dependencies = TRUE)
}
if (!is.element("ggcorrplot", installed.packages()[, 1])) {
  install.packages("ggcorrplot", dependencies = TRUE)
}
library(corrplot)
library(ggcorrplot)

# Create correlation plots for your datasets

# Using corrplot
corrplot(cor(Student_Perfomance_Dataset[, sapply(Student_Perfomance_Dataset, is.numeric)]), method = "circle")

# Using ggcorrplot for a visually appealing plot
ggcorrplot(cor(Student_Perfomance_Dataset[, sapply(Student_Perfomance_Dataset, is.numeric)]))

```

```{r code chunk - Used to create a scatter plot - step 22}

if (!is.element("ggplot2", installed.packages()[, 1])) {
  install.packages("ggplot2", dependencies = TRUE)
}
library(ggplot2)

ggplot(Student_Perfomance_Dataset, aes(x = `Absenteeism Percentage`, y = GRADE)) +
  geom_point() + geom_smooth(method = lm)

```

```{r code chunk - Used to create Multivariate Box and Whisker Plots by Class - step 23}

# Execute the following code:
if (!is.element("caret", installed.packages()[, 1])) {
  install.packages("caret", dependencies = TRUE)
}
require("caret")

featurePlot(x = Student_Perfomance_Dataset[, 23], y = Student_Perfomance_Dataset[, 26], plot = "box")

```
