---
title: "Business Intelligence Lab Submission Markdown"
author: "<Specify your name here>"
date: "<Specify the date when you submitted the lab>"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

+---------------------------------------------------+--------------------------------------------------------------------------------------------+
| **Student ID Numbers and Names of Group Members** | \<list one student name and ID per line; you should be between 2 and 5 members per group\> |
|                                                   |                                                                                            |
|                                                   | 1.122998-Starfield-Glenn Oloo                                                              |
|                                                   |                                                                                            |
|                                                   | 2.134782-Starfield-Yasmin Choma                                                            |
|                                                   |                                                                                            |
|                                                   | 3.134783-Starfield- Moses Mbugua                                                           |
+---------------------------------------------------+--------------------------------------------------------------------------------------------+
| **BBIT 4.2 Group**                                | \<either group A, B, or C\>                                                                |
+---------------------------------------------------+--------------------------------------------------------------------------------------------+
| **Course Code**                                   | BBT4206                                                                                    |
+---------------------------------------------------+--------------------------------------------------------------------------------------------+
| **Course Name**                                   | Business Intelligence II                                                                   |
+---------------------------------------------------+--------------------------------------------------------------------------------------------+
| **Program**                                       | Bachelor of Business Information Technology                                                |
+---------------------------------------------------+--------------------------------------------------------------------------------------------+
| **Semester Duration**                             | 21^st^ August 2023 to 28^th^ November 2023                                                 |
+---------------------------------------------------+--------------------------------------------------------------------------------------------+

# Setup Chunk

**Note:** the following KnitR options have been set as the global defaults:\
`knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE,                        collapse = FALSE, tidy = TRUE)`.

> output:\
> \
> github_document:\
> toc: yes\
> toc_depth: 4\
> fig_width: 6\
> fig_height: 4\
> df_print: default\
> \
> editor_options:\
> chunk_output_type: console

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE,
                      collapse = FALSE, tidy = TRUE)
```

# \<Provide an Appropriate Title Here\>

Describe the code chunk here:

```{r Your Second Code Chunk -Code used to load Student Perfomance dataset}
library(readxl)
Student_Perfomance_Dataset <- read_excel("data/Student Perfomance Dataset.xlsx")
View(Student_Perfomance_Dataset)
```

```{r Your Second Code Chunk Used to check the number of Variables and Rows in the student Perfomance Dataset}
library(readr)
dim(Student_Perfomance_Dataset)
```

Describe the next code chunk here:

```{r Your Third Code Chunk- Used to check the data types of the Variables which can either be Numeric or Character}
library(readr)
sapply(Student_Perfomance_Dataset, class)
```

## \<You Can Have a Sub-Title Here\>

```{r Your Fourth Code Chunk - used to see the frequency for the variables that are categorical i.e. See the frequency of students that have mentors and those who don't}

library(readr)
Student_Perfomance_Dataset_freq <- Student_Perfomance_Dataset$mentor
cbind(frequency = table(Student_Perfomance_Dataset_freq),
      percentage = prop.table(table(Student_Perfomance_Dataset_freq)) * 100)

Student_Perfomance_Dataset_freq <- Student_Perfomance_Dataset$`Attendance Waiver Granted: 1 = Yes, 0 = No`
cbind(frequency = table(Student_Perfomance_Dataset_freq),
      percentage = prop.table(table(Student_Perfomance_Dataset_freq)) * 100)
```


```{r Your Third Code Chunk- Used to perform the mode, that is the variable that is repeated the most in Mentor "0" is the mode meaning most students do not have mentors}
library(readr)
Student_Perfomance_Dataset_mode <- names(table(Student_Perfomance_Dataset$mentor))[
  which(table(Student_Perfomance_Dataset$mentor) == max(table(Student_Perfomance_Dataset$mentor)))
]
print(Student_Perfomance_Dataset_mode)

Student_Perfomance_Dataset_mode <- names(table(Student_Perfomance_Dataset$`Attendance Waiver Granted: 1 = Yes, 0 = No`))[
  which(table(Student_Perfomance_Dataset$`Attendance Waiver Granted: 1 = Yes, 0 = No`) == max(table(Student_Perfomance_Dataset$`Attendance Waiver Granted: 1 = Yes, 0 = No`)))
]
print(Student_Perfomance_Dataset_mode)

```


```{r Your Third Code Chunk- Used to calculate the distribution of each variable that is Median, mean , 1st Quartile,3rd Quartile}
library(readr)
summary(Student_Perfomance_Dataset)
```


```{r Your Third Code Chunk- Used to calculate the the standard deviation for variables that data type are numeric}
library(readr)

numeric_columns <- Student_Perfomance_Dataset[, sapply(Student_Perfomance_Dataset, is.numeric)]
standard_deviations <- sapply(numeric_columns, sd)
print(standard_deviations)
```

```{r Your Third Code Chunk- Used to calculate the variance of each variable}
library(readr)
numeric_columns <- Student_Perfomance_Dataset[, sapply(Student_Perfomance_Dataset, is.numeric)]
variances <- sapply(numeric_columns, var)
print(variances)

```
**etc.** as per the lab submission requirements.
